---
title: "Milestone 3"
output: html_document
date: "2023-10-17"
---

```{r}
library(statsr)
library(dplyr)
library(ggplot2)
library(readr)
library(multcomp)
```


```{r}
setwd("/Users/oscarfs/GitHub/SQL-Capstone_Project")
```


```{r}
df <- read_csv("export.csv")
df$Medal[is.na(df$Medal)] <- "No medal"    # Change NA's to No medal.
df$Medal <- as.factor(df$Medal)

summary(df)

```

## Effect of Age on performance

From the previous milestone, we remember that we saw some differences between the different classes with age. The next step is to compare them and find if the differences are significant.

```{r}
ggplot(data = df, aes(x = Medal, y = Age)) + geom_boxplot()
```

These are the medias of the age per group.

```{r}
df %>% group_by(df$Medal) %>%
  summarise(mean = mean(Age), SD = sd(Age))
```

We will conduct an Analysis of Variance (ANOVA) since we have more than two means. We also need to adjust the $\alpha$ value with the Bonferroni correction. So the new p_value is calculated as $\alpha / K$, where $K = \dfrac{k(k-1)}{2}$. So, the new p-value is:

```{r}
# New p-value.
0.05/(4*3/2)
```

The results of the ANOVA test are the following:

```{r}
inference(y = Age, x = Medal, data = df, statistic = "mean", type = "ht", conf_level = 0.95, null = 0, alternative = "greater", method = "theoretical", order = c("Gold", "Silver", "Bronze", "No medal"))
```

From these results we can notice that the p-value is less than 0.0001. Therefore, we can conclude that there is at least one pair of means where the difference is significant.  

The pairs where this difference is significant ($p\_value < 8.33\text{x}10^{-3}$) are:
* No medal vs Gold: Athletes with no medals have a higher age than top performers (athletes that earn gold medals).
* No medal vs Bronze: Athletes who won bronze medals are younger than athletes with no medals.
* Silver vs No medal: Interestingly, athletes that perform better than those with bronze and no medals, are older than those who perform worst.

Therefore, age might play a role in the probability of performing better in any sport.


### Simplifying the performance metric

We now have two classes: Medalist or not medalist.

```{r}
df2 <- df
df2$Medal = ifelse(df$Medal == 'No medal', "No", "Yes")
df2$Medal = as.factor(df2$Medal)

summary(df2)
```

The differences in the mean are slight, but still we can see that medalist are slightly younger.

```{r}
df2 %>% group_by(Medal) %>%
  summarise(mean = mean(Age), SD = sd(Age))
```

The difference is significant with p-value less than 0.0001. Therefore, younger athletes tend to perform better than older athletes.

```{r}
inference(y = Age, x = Medal, data = df2, statistic = "mean", type = "ht", null = 0, alternative = "greater", method = "theoretical")
```


## New metric: Body Mass Index (BMI)

We can see that Height and Weight have a strong positive correlation. We decided to join both attributes into a new metric that shows the meassure of both: the BMI. This is calculated as $BMI = \dfrac{Weight}{Height^2}$. The units are km/m$^2$.

```{r}
ggplot(data = df, aes(x = Height, y = Weight)) +
  geom_point(size = 0.2) +
  geom_smooth(method = "lm") +
  annotate("text",
    x = 160, y = 150,
    label = paste("Correlation = ", round(cor(df$Height, df$Weight, method = "pearson"), 4)))
```

```{r}
df <- df %>%
    mutate(BMI = Weight / (Height / 100) ** 2)

head(df, n = 10)
```

There is no strong correlation between Age and the new metric. The higher the performance, the higher the correlation, but this is not very strong.

```{r}
aux_df <- df[df$Medal == "No medal", ]
ggplot(data = aux_df, aes(x = Age, y = BMI)) +
  geom_point(size = 0.2) +
  geom_smooth(method = "lm") +
  annotate("text",
    x = 25, y = 45,
    label = paste("No medal correlation = ", round(cor(aux_df$Age, aux_df$BMI, method = "pearson"), 4)))

```

```{r}
aux_df = df[df$Medal == "Bronze", ]
ggplot(data = aux_df, aes(x = Age, y = BMI)) +
  geom_point(size = 0.2) +
  geom_smooth(method = "lm") +
  annotate("text",
    x = 25, y = 45,
    label = paste("Bronze correlation = ", round(cor(aux_df$Age, aux_df$BMI, method = "pearson"), 4)))
```

```{r}
aux_df = df[df$Medal == "Silver", ]
ggplot(data = aux_df, aes(x = Age, y = BMI)) +
  geom_point(size = 0.2) +
  geom_smooth(method = "lm") +
  annotate("text",
    x = 25, y = 45,
    label = paste("Silver correlation = ", round(cor(aux_df$Age, aux_df$BMI, method = "pearson"), 6)))
```

```{r}
aux_df = df[df$Medal == "Gold", ]
ggplot(data = aux_df, aes(x = Age, y = BMI)) +
  geom_point(size = 0.2) +
  geom_smooth(method = "lm") +
  annotate("text",
    x = 25, y = 45,
    label = paste("Gold correlation = ", round(cor(aux_df$Age, aux_df$BMI, method = "pearson"), 4)))
```

With these new metric, there seems to be a difference in the BMI for each performance group.

```{r}
ggplot(data = df, aes(x = Medal, y = BMI)) + geom_boxplot()
```
It seems that top athletes have a higher BMI than athletes without medals.

```{r}
df %>% group_by(Medal) %>%
  summarise(mean = mean(BMI), SD = sd(BMI))
```

To find significant differences between these means,we are interested in the ANOVA test for the new metric.

```{r}
inference(y = BMI, x = Medal, data = df, statistic = "mean", type = "ht", conf_level = 0.95, null = 0, alternative = "greater", method = "theoretical", order = c("Gold", "Silver", "Bronze", "No medal"))
```

The results show that the p-value is less than 0.001. Meaning that there is at least one pair of means where the difference is significant. These groups are the following:
* No medal vs Gold: Top athletes have a lower IBM than athletes with the lowest performance.
* No medal vs Bronze: Athletes with lower performance have lower IBM than those with Bronze medals.
* Silver vs No medal: Silver medal athletes have higher IBM than the athletes with the lowest performance.


```{r}
res_aov <- aov(BMI ~ Medal,
  data = df
)

summary(res_aov)
```

```{r}
post_test <- glht(res_aov,
  linfct = mcp(Medal = "Tukey")
)

summary(post_test)
```

### Using two-tier performance metric

```{r}
df2 <- df
df2$Medal = ifelse(df$Medal == 'No medal', "No", "Yes")
df2$Medal = as.factor(df2$Medal)

summary(df2)
```

```{r}
df2 %>% group_by(Medal) %>%
  summarise(mean = mean(BMI), SD = sd(BMI))
```

```{r}
inference(y = BMI, x = Medal, data = df2, statistic = "mean", type = "ht", null = 0, alternative = "greater", method = "theoretical", order = c("Yes", "No"))
```


## BMI across different sports

```{r}
ggplot(df, aes(x = Sport, y = BMI)) + geom_boxplot()
```


```{r}
inference(y = BMI, x = Sport, data = df, statistic = "mean", type = "ht", conf_level = 0.95, null = 0, alternative = "greater", method = "theoretical")
```







